### 2013-09-16 chi2014/future.tex

\section{Future Work}

\subsection{Interventions}

\begin{itemize}
\item multi-phase brainstorming: introducing common ideas first to force workers more quickly into the originality zone
\end{itemize}

\subsection{Predicting Turker performance}

\subsection{Automatic Idea Forest Generation}